---
title: "Assess Depth"
output: html_document
---

Purpose: to assess the coverage depth at our PdMyb1 locus and find a window 
within which we can trust SNP calls and perform population genetic analyses.

# Load packages
```{r loadPkg, message=FALSE}
# Install packages
req_pkgs <- c("data.table", "stringr", "ggpubr", "patchwork")
for(pkg in req_pkgs){
    if(!require(pkg, character.only=TRUE)){install.packages(pkg, repos="https://cloud.r-project.org/"); library(pkg, character.only=TRUE)}
}
```

# Load data
## Gene annotation data
```{r loadAnnot}
gene_d <- fread("annot/PdMyb1.gene.bed", header=FALSE, col.names=c("CHROM", "START", "END", "ELEMENT"), sep="\t")
print(gene_d)
```

## VCF data

First we use bcftools query to grab the mpileup depth values

```{bash runQuery}
# bash chunk

for sp in drum cusp; do
    merge_vcf=results/bcf_pipe/call/merged.${sp}.vcf.gz
    qry_fn=${merge_vcf%.vcf*}.DP.qry
    if [ ! -s $qry_fn ]; then
        bcftools query -Hf '%CHROM\t%POS[\t%DP]\n' $merge_vcf > $qry_fn
    fi
done
```

Now we load in that data
```{r loadQuery}
depth_d <- list()
for(sp in c("drum", "cusp")){
    tmp <- fread(str_interp("results/bcf_pipe/call/merged.${sp}.DP.qry"), header=TRUE, sep="\t", 
                    integer64="numeric", na.strings=".")
    names(tmp)[1] <- str_remove(names(tmp)[1], "^#\\[1\\]")
    names(tmp) <- str_remove(names(tmp), "^\\[[0-9]+\\]")
    names(tmp) <- str_remove(names(tmp), ":DP$")
    depth_d[[sp]] <- tmp
    rm(tmp); gc()
}
```

## Sample names
```{r loadSamp}
all_samps <- list()
for(sp in names(depth_d)){
    all_samps[[sp]] <- fread(str_interp("config/${sp}.samples.list"), header=FALSE, sep="\t")$V1
    if(!all(all_samps[[sp]] %in% names(depth_d[[sp]]))){
        stop("Not all samples represented in depth data")
    }
}
```

# Aggregate depth
Aggregate depth by taking mean depth in 1 kb windows.

NB: We assume that sites with NA depth have 0 coverage

First bin/assign positions into windows

```{r binPos}
wsize <- 1000

pos_range <- depth_d[["drum"]][, range(POS)]
pos_range[1] <- floor(pos_range[1] / wsize ) * wsize
pos_range[2] <- ceiling(pos_range[2] / wsize ) * wsize


# Bin positions every 100 0
aggr_breaks <- seq(pos_range[1], pos_range[2], by=wsize)
for(sp in names(depth_d)){
    depth_d[[sp]][, POS_BIN := cut(POS, aggr_breaks, labels=aggr_breaks[-length(aggr_breaks)])]
    depth_d[[sp]][, POS_BIN := as.numeric(as.character(POS_BIN))]
}
```

Next do the actual averaging.

```{r aggrDepth}
aggr_d <- list()
for(sp in names(depth_d)){
    for(samp in all_samps[[sp]]){
        depth_d[[sp]][is.na(get(samp)), eval(samp) := 0]
        aggr_d[[samp]] <- depth_d[[sp]][, .(MEAN_DP = mean(get(samp))), by=POS_BIN]
        aggr_d[[samp]][, SAMPLE := samp]
        aggr_d[[samp]][, SPECIES := sp]
    }
}
aggr_d <- rbindlist(aggr_d)
print(head(aggr_d))
```

# Plot!

Plot the depth dropoff around the gene

```{r plotDepth, fig.width=8, fig.height=4}
AnnotateExons <- function(p, y=-10, lw=3){
    use_color <- "#666666"

    exonrange <- gene_d[str_starts(ELEMENT, "exon"), range(c(START, END))/1e3]
    p <- p + annotate(geom="segment", x=exonrange[1], xend=exonrange[2], 
                    y=y, yend=y, color=use_color, size=lw/7)

    for(i in 1:nrow(gene_d)){
        if(str_starts(gene_d[i, ELEMENT], "exon")){
            p <- p + annotate(geom="segment", x=gene_d[i, START/1e3], xend=gene_d[i, END/1e3], 
                            y=y, yend=y, color=use_color, size=lw)
        }
    }
    return(p)
}


use_xlim <- c(-50, 50) + median(range(gene_d[, c(START, END)]))/1e3

# Median and interquartile range across samples
med_dp <- aggr_d[, as.list(quantile(MEAN_DP, probs=c(0.25, 0.5, 0.75))), by=.(POS_BIN)]
setnames(med_dp, c("25%", "50%", "75%"), c("LWR", "MED", "UPR"))


p_base <- ggplot() +
    geom_line(aes(x=POS_BIN/1e3, y=MEAN_DP, group=SAMPLE), aggr_d[SPECIES=="drum"], alpha=0.1, color="black") +
    geom_line(aes(x=POS_BIN/1e3, y=MEAN_DP, group=SAMPLE), aggr_d[SPECIES=="cusp"], color="coral", alpha=0.8) +
    geom_ribbon(aes(x=POS_BIN/1e3, ymin=LWR, ymax=UPR), med_dp, color=NA, fill="dodgerblue", alpha=0.25) +
    geom_line(aes(x=POS_BIN/1e3, y=MED), med_dp, color="dodgerblue4", alpha=1.0, size=0.3)


p_base <- p_base + xlab("Position (kb)") + ylab("Mean depth (1 kb bins)") +
    theme_pubr() + scale_x_reverse()

p_depth <- AnnotateExons(p_base, y=-10, lw=2.5) + coord_cartesian(xlim=rev(use_xlim), ylim=c(0,500))

print(p_depth)
```

Based on observing the general pattern across samples (and messing around with with plotly, not shown), settling on 
the window: 

`chrlg6A 22289000 22324000`

```{r plotWithWindow, fig.width=12, fig.height=8}
wstart <- 22289000L
wend   <- 22324000L

p_base <- p_base + 
    geom_vline(xintercept=wstart/1e3, linetype="dotted") +
    geom_vline(xintercept=wend/1e3,   linetype="dotted")

p_A <- AnnotateExons(p_base, y=-20, lw=3) + coord_cartesian(xlim=rev(use_xlim))                + ggtitle("A") + theme(plot.title=element_text(size=18))
p_B <- AnnotateExons(p_base, y=-10, lw=3) + coord_cartesian(xlim=rev(use_xlim), ylim=c(0,500)) + ggtitle("B") + theme(plot.title=element_text(size=18))

p_out  <- p_A / p_B + plot_layout(heights=c(1, 1))

ggsave(p_out, file="figureS6.pdf", width=12, height=8)

print(p_out)
```

Save the final window as a bed file
```{r saveBed}
out_d <- data.table(CHROM="chrlg6A", START=wstart, END=wend)

out_fn <- "results/bcf_pipe/call/drum.hi_dep_region.bed"

fwrite(out_d, file=out_fn, sep="\t", row.names=FALSE, col.names=FALSE, quote=FALSE)
```

