---
title: "GWAS Plotting"
output: html_document
---


## Load packages
```{r loadPkg, message=FALSE}
# Install packages
req_pkgs <- c("data.table", "stringr", "ggpubr")
for(pkg in req_pkgs){
    if(!require(pkg, character.only=TRUE)){install.packages(pkg, repos="https://cloud.r-project.org/"); library(pkg, character.only=TRUE)}
}
```

## Load data
```{r loadRes}
gwas_prefix <- "results/gwas/merged.drum.filt.lmiss_0.5.imiss_0.9.polarized.bi"
pheno_nms <- c("Qual_Int", "Quant_Chroma", "Quant_Int_Bright")
dat <- list()
for(pheno in pheno_nms){
    # Note that Brightness is essentially the quantitative trait
    # for Intensity hence the name "Quant_Int_Bright"
    fn <- str_interp("${gwas_prefix}.${pheno}.gemma.relmat.maf_0.025.ulmm.assoc.txt")
    dat[[pheno]] <- fread(fn, header=TRUE)
    dat[[pheno]][, PHENO := pheno]
}
dat <- rbindlist(dat)
dat[, COORD := as.numeric(ps) / 1e3]
```

## Identify outliers

First, let's calculate the bonferroni threshold (alpha = 0.05)

```{r bonferroni}
a <- 0.05
bonf_thresh <- a / nrow(dat[PHENO == "Qual_Int"])
cat(str_interp("a = $[0.02f]{a}\n"))
cat(str_interp("-log10( a_corrected ) = $[0.1f]{-log10(bonf_thresh)} \n"))
```

Now, report all SNPs below threshold
```{r}
print(dat[p_wald < bonf_thresh, ])
```

For all phenotypes, we have two snps below the threshold:
- SNP1: 22302338
- SNP2: 22302416
```{r defineOutliers}
outlier_snps <- dat[PHENO == "Qual_Int"][p_wald < bonf_thresh, ps[order(p_wald)]]
cat(str_interp("$[d]{outlier_snps[1]}\n"), file="annot/PdMyb1.putative_snp.pos")
```

## Report log p values 
```{r reportPvals}
for(i in seq_along(outlier_snps)){
    snp <- outlier_snps[i]
    cat(str_interp("SNP${i}, pos = $[d]{snp}:\n"))
    print(dat[ps == snp, .(PHENO, LOG10P=-log10(p_wald))])
    cat("\n")
}
```
## PVE: Regress phenotypes against outlier SNPs

First we need genotypes at the SNPs
```{r loadGTs}
snp_gts <- list()
for(pheno in pheno_nms){ 
    tmp <- fread(str_interp("${gwas_prefix}.${pheno}.recode.traw"), header=TRUE, sep="\t", na.strings="NA")[POS %in% outlier_snps]

    snp_gts[[pheno]] <- fread(str_interp("${gwas_prefix}.${pheno}.fam"), header=FALSE, col.names=c("FID", "IID", "PAT_ID", "MAT_ID", "SEX", "VAL"))
    snp_gts[[pheno]][, PHENO := pheno]
    fid_iid <- snp_gts[[pheno]][, str_c(FID, "_", IID)]
    
    for(i in seq_along(outlier_snps)){
        new_col <- str_c("SNP", i)
        new_val <- unlist(tmp[POS == outlier_snps[i], fid_iid, with=FALSE])
        snp_gts[[pheno]][, eval(new_col) := new_val]
    }
}

snp_gts <- rbindlist(snp_gts)
```

Now let's regress to get percent variance explained by each snp (PVE)

```{r regression}
calcPVE <- function(in_dat, xvar="SNP1", yvar="VAL"){
    fmla <- as.formula(str_interp("${yvar} ~ ${xvar}"))
    m <- lm(fmla, in_dat)
    return(summary(m)$r.squared)
}

```


## Load LD data at SNP1

Let's load the R^2 results for SNP1 pairwise to all other snps.

```{r loadLD}
ld_prefix <- "results/pairwise_ld/merged.drum.filt.lmiss_0.5.imiss_0.9.polarized.bi"
ld_pos <- fread(str_interp("${ld_prefix}.bim"), header=FALSE)$V4
keep_col <- which(ld_pos == outlier_snps[1])

ld_vals <- fread(str_interp("${ld_prefix}_ld.ld"), header=FALSE, select=keep_col)
names(ld_vals) <- "RSQ"
ld_vals[, ps := ld_pos]
```

What's R^2 between SNP1 and SNP2?

```{r reportLD}
x <- ld_vals[ps == outlier_snps[2], RSQ]
cat(str_interp("R^2 between SNP1 and SNP2: $[0.2f]{x}\n"))
```

## Plot results

Need a function for transforming the -log10 transformation on the y-axis.

```{r plotFuncs}
neg_log10_trans <- function() {
    scales::trans_new(
        name = "neg_log10",
        transform = function(x) -log10(x),
        inverse = function(x) 10^(-x)
    )
}
```


Manhattan plot

```{r plotManhat, fig.width=12, fig.height=4}
# Add R2 information
dat <- merge(dat, ld_vals, by="ps", all.x=TRUE, all.y=FALSE)

for(pheno in pheno_nms){
    pwrs <- seq(-2, -18, by=-4)
    brks <- 10 ^ pwrs
    labs <- sprintf("%d", -pwrs)

    p <- ggscatter(dat[PHENO == pheno], x="COORD", y="p_wald", alpha=0.5, color="RSQ") + 
        scale_y_continuous(trans=neg_log10_trans(), breaks=brks, labels=labs) +
        ylab(bquote(- log[10](p))) +
        scale_x_reverse() +
        xlab("chr6 position (kb)") +
        ggtitle(pheno)
    print(p)
}
# ggsave(p, file="tmp.pdf", width=12, height=4)
```
